An exploration of Wine Quality  by James Cooper
========================================================

# Introduction

This report will explore the White Wine dataset and see what conclusions we can
draw about what makes a good wine.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(knitr)
library(dplyr)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
getwd()
wd <- read.csv('C:/Users/James2SxyBoogaloo/Google Drive/Data/P4 - EDA/wineQualityReds.csv', header = TRUE)
```
# Univariate Plots Section

## Distribution of Each Variable

FIXED ACIDITY:

```{r echo=FALSE, Univariate_Plot_var2}
plot.hist <- function(x.var, wd, bw){
  ggplot(wd, aes_string(x = x.var)) +
    geom_histogram(binwidth = bw)
}

plot.hist('fixed.acidity', wd, 0.1)
#ggplot(data = wd,
      # aes(x = fixed.acidity)) +
       #geom_histogram(binwidth = 0.1)
```

The values appear to be clustered around 8 however let's do a simple summary to
see what we get. I'll also include a boxplot to represent this visually.

```{r echo=FALSE, Univariate_Plot_var2_sum}
summary(wd$fixed.acidity)
```

```{r echo=FALSE, Univariate_Plots_var2_2}
ggplot(data = wd,
       aes(x = X, y = fixed.acidity)) +
       geom_boxplot()
```

There are a few outliers that have nearly double fixed acidity of the other
wines.

I will use summary tables instead of boxplots going forward, unless there is a
particular reason not to.

VOLATILE ACIDITY:

```{r echo=FALSE, Univariate_Plots_var3}
plot.hist('volatile.acidity', wd, 0.01)
```
```{r echo=FALSE, Univariate_Plot_var3_sum}
summary(wd$volatile.acidity)
```

IQR of 0.39 to 0.64

Outliers - A few on the high side. 

CITRIC ACID:

```{r echo=FALSE, Univariate_Plots_var13}
plot.hist('citric.acid', wd, 0.01)
```

```{r echo=FALSE, Univariate_Plot_var13_sum}
summary(wd$citric.acid)
```

IQR of 0.09 to 1

Outliers - A single high outlier

RESIDUAL SUGAR:
```{r echo=FALSE, Univariate_Plots_var4}
plot.hist('residual.sugar', wd, 0.3)
```

```{r echo=FALSE, Univariate_Plot_var4_sum}
summary(wd$residual.sugar)
```

Distance between the third quartile and the max is massive - several high
outliers. This meant the data was a good fit for the use of a transformation.

CHLORIDES:

```{r echo=FALSE, Univariate_Plots_var5}
plot.hist('chlorides', wd, 0.01)
```

Chloride data was heavily skewed so I used log10 once again - we can more
clearly see the fairly normal distribution here.

```{r echo=FALSE, Univariate_Plot_var5_sum}
summary(wd$chlorides)
```

Again, large difference between the 3rd quartile and max. A few values on the
low end as well - whilst visually these are not outliers, it could be worth
keeping track of these results.

SULFUR DIOXIDE:
```{r echo=FALSE, Univariate_Plots_var6}
plot.hist('free.sulfur.dioxide', wd, 2)
```

```{r echo=FALSE, Univariate_Plot_var6_sum}
summary(wd$free.sulfur.dioxide)
```

Yet again we can see that several outliers are present, and the familiar
pattern of a large top quartile appears once again. The distribution is skewed,
so lets see what taking the log does:

```{r echo=FALSE, Univariate_Plots_var66}
ggplot(data = wd,
       aes(x = free.sulfur.dioxide)) +
       geom_histogram(binwidth =0.05) +
scale_x_log10()
```

The log adjusted FS02 figures show a bimodal distribution.

TOTAL SULFUR DIOXIDE:

```{r echo=FALSE, Univariate_Plots_var7}
plot.hist('total.sulfur.dioxide', wd, bw = 3)
```

```{r echo=FALSE, Univariate_Plot_var7_sum}
summary(wd$total.sulfur.dioxide)
```

As above.

```{r echo=FALSE, Univariate_Plots_var76}
ggplot(data = wd,
       aes(x = total.sulfur.dioxide)) +
       geom_histogram() +
  scale_x_log10()
```

Taking the log reveals a nice normal distribution.

DENSITY:
```{r echo=FALSE, Univariate_Plots_var8}
plot.hist('density', wd, 0.0002)

```

```{r echo=FALSE, Univariate_Plot_var8_sum}
summary(wd$density)
```

A normal distribution - however some of the wines appear to be less dense than
water! This is odd because we would expect density to increase because of sugar
and alcohol content.

pH:

```{r echo=FALSE, Univariate_Plots_var9}
plot.hist('pH', wd, 0.01)
```

```{r echo=FALSE, Univariate_Plot_var9_sum}
summary(wd$pH)
```

A fairly well spread out distribution. As pH is logarithmic, a difference of 1
tells us that there are actually some wines that are more than ten times as
acidic as others!

SULPHATES:

```{r echo=FALSE, Univariate_Plots_var10}
plot.hist('sulphates', wd, 0.01)
```

```{r echo=FALSE, Univariate_Plot_var10_sum}
summary(wd$sulphates)
```

A high Q4 again - and some significant outliers.

ALCOHOL:

```{r echo=FALSE, Univariate_Plots_var11}
plot.hist('alcohol', wd, 0.1)

```

```{r echo=FALSE, Univariate_Plot_var11_sum}
summary(wd$alcohol)
```

QUALITY:

```{r echo=FALSE, Univariate_Plots_var12}
ggplot(data = wd,
       aes(x = quality)) +
       geom_bar()
```

```{r echo=FALSE, Univariate_Plot_var12_sum}
summary(wd$quality)
```

# Univariate Analysis

STRUCTURE:

```{r echo=FALSE, STR}
str(wd)
```

### What is the structure of your dataset?

The following information taken directly from 
https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt

 Description of Input Variables

   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do
   not evaporate readily)

   2 - volatile acidity: the amount of acetic acid in wine, which at too high
   of levels can lead to an unpleasant, vinegar taste

   3 - citric acid: found in small quantities, citric acid can add 'freshness' 
   and flavor to wines

   4 - residual sugar: the amount of sugar remaining after fermentation stops, 
   it's rare to find wines with less than 1 gram/liter and wines with greater
   than 45 grams/liter are considered sweet

   5 - chlorides: the amount of salt in the wine

   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between 
   molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial
   growth and the oxidation of wine

   7 - total sulfur dioxide: amount of free and bound forms of S02; in low 
   concentrations, SO2 is mostly undetectable in wine, but at free SO2
   concentrations over 50 ppm, SO2 becomes evident in the nose and taste of
   wine

   8 - density: the density of water is close to that of water depending on the
   percent alcohol and sugar content

   9 - pH: describes how acidic or basic a wine is on a scale from 0
   (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas
   (S02) levels, wich acts as an antimicrobial and antioxidant

   11 - alcohol: the percent alcohol content of the wine

   Output variable (based on sensory data): 
   
   12 - quality (score between 0 and 10)

An interesting note is that the authors point out that they are unsure whether
all the input variable are relevant. If too many variables with a high 
correlation are left in this could harm the regression.

### What is/are the main feature(s) of interest in your dataset?

Main feature of interest is the quality scores as this is the output variable.
The main point of this exploration is to find out which input variables are
useful predictors of quality.This is of course subject to the major caveat that
Quality itself is not an objective measure, and furthermore that the quality of 
this non- objective measure (a wine experts ability to consistently and 
accurately grade wine in a controlled blind environment) has been called into
question (read: completely debunked) by numerous recent studies.

www.theguardian.com/lifeandstyle/2013/jun/23/wine-tasting-junk-science-analysis

### What other features in the dataset do you think will help support your \
###investigation into your feature(s) of interest?

I think that the outliers will provide a useful insight into the grades
received. I would expect that the wines with a massive concentration of salt /
acid would tend towards lower scores.

### Did you create any new variables from existing variables in the dataset?

No - I felt that whilst this woud be possible ( IE create bound S02 measure by
subtracting free SO2 from total S02), I felt that this was unlikely to provide
any extra insight. After all, the curators have not included it in the dataset
- I am assuming they know more about wine than me.

### Of the features you investigated, were there any unusual distributions?

Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

No - as mentioned above, I want to retain the outliers for now. Later I can
exclude them if necessary. However in the plots above I have also included more
zoomed in graphs that better show the bulk of the data.

# Bivariate Plots Section

QUALITY VERSUS FREE SO2

```{r echo=FALSE, Bivariate_Plot1}
ggplot(aes(x = quality, y = free.sulfur.dioxide),
             data = wd) +
  geom_jitter(alpha = 0.4)
```

This doesn't really show a clear relationship - there are so many more
observations for quality 5 and 6 wines that it is difficult to get a feel for
the data.

QUALITY VERSUS Total SO2

```{r echo=FALSE, Bivariate_Plot2}
ggplot(aes(x = quality, y = total.sulfur.dioxide),
             data = wd) +
  geom_jitter(alpha = 0.4)
```

THese graphs don't really show us too much however of some note are the 2
outliers that got a 7 in taste despite havingthe highest So2 scores. THis seems 
odd!

REVISITNG Q VS FREE SO2 and Q vs bound SO2 - this time  using boxplots.

QUALITY VERSUS free SO2 2

```{r echo=FALSE, Bivariate_Plot3}

ggplot(aes(x = quality, y = free.sulfur.dioxide, group = quality),
             data = wd) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 50))

```

Quality versus total So2 2

```{r echo=FALSE, Bivariate_Plot4}
ggplot(aes(x = quality, y = total.sulfur.dioxide, group = quality),
             data = wd) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 175))
  
```

Note the mean similarities.

```{r echo=FALSE, x}
with(wd, cor.test(free.sulfur.dioxide, quality, method = 'pearson'))

with(wd, cor.test(total.sulfur.dioxide, quality, method = 'pearson'))

```

There is a relationship here, but it is extremely weak. This means that 1. 
Outliers are masking due to experimenal error or 2. other factors are masking
this due to interrelationships between variables.

We can at least address the first one by removing the outliers. Let's see what
effect that has.There are mutiple ways to do this. I will simply exclude FSO
over 60.

```{r echo=FALSE, xs}
with(subset(wd, free.sulfur.dioxide <= 60), cor.test(free.sulfur.dioxide, quality))
```

Excluding outliers actually reduces the correlation for Fs02.

And for Total FS02:

```{r echo=FALSE, xss}
with(subset(wd, total.sulfur.dioxide <= 200), cor.test(total.sulfur.dioxide, quality))
```

On the other hand, exluding the outliers actually strengthens the relationship
between TSO2 and quality.

THis is a bit confusing - the wording used in the model implied that there would
be a negative impact on the taste of the wine based on the free rather than
total sulphur content.
However there are several other variables at work here and it is possible that
this is simply not a major factor relative to them. Before giving up on this
line of inquiry I will attempt to examine this one last time.

I am going to take an average of the Quality score for each concentration,
flipping the chart in the process.

I will use Dplyr to group by free.sulfur.dioxide.

```{r echo=FALSE, anulook}
library(dplyr)
free_groups <- group_by(wd, free.sulfur.dioxide)

wd.qual_by_free <- summarise(free_groups,
                             quality_mean = mean(quality),
                             quality_median = median(quality),
                             n = n())
head(wd.qual_by_free)
```
```{r echo=FALSE, anulook2}
ggplot(aes(x = free.sulfur.dioxide, y = quality_mean),
       data = wd.qual_by_free) +
  geom_point() +
  geom_smooth()
```

THere is a lot of variance introduced by individual readings.
If we round FSO we can smooth some of this out to see if the relationship 
becomes a little clearer.

```{r echo=FALSE, anulook26}
ggplot(aes(x = round(free.sulfur.dioxide / 2) * 2, y = quality_mean),
       data = wd.qual_by_free) +
  geom_point() +
  geom_smooth()
```

Unfortunately this approach does not really give us a smoother line as expected.

```{r echo=FALSE, anulook3}
ggplot(aes(x = free.sulfur.dioxide, y = n),
       data = wd.qual_by_free) +
  geom_point()
```

There is also a massive difference in the number of observations for each
concentration. Clearly the means for the quality of the higher concentrations of
SO2 are going to be much less accurate than those on the lower end.

```{r echo=FALSE, bins}
wd.qual_by_free$fs02bin <- as.numeric(cut_number(wd.qual_by_free$free.sulfur.dioxide, 200))

ggplot(aes(x = fs02bin, y = quality_mean),
       data = wd.qual_by_free) +
  geom_point() +
  geom_smooth()

```

This was the first attempt to bin FS02 - it failed.

I will come back to this graph as it raises some interesting questions.

For now I will end the exploration of FS02 by itself by examining the median
rather than the mean.

```{r echo=FALSE, bins6}
ggplot(aes(x = fs02bin, y = quality_median),
       data = wd.qual_by_free) +
  geom_point() +
  geom_smooth()

```

I should have known that this would not provide any useful insight.

Let's take a step back and see how TSO2 and FSO2 are related:

```{r echo=FALSE, SUlphur}
ggplot(data = wd,
       aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
      geom_point(alpha = 0.2,
                 position = position_jitter(h = 0),
                 color = 'blue') +
      geom_line(stat = 'summary', fun.y = mean) +
      geom_smooth(method = 'lm', color = 'red')
```

One interesting thing about the boxplots further above is that the free and
total S02 appear to show similar distributions, if we look at the mean for each
quality level.This appears to be consistent with the fact that they are
moderately correlated with each other, as demonstrated here.

Corr to confirm:

```{r echo=FALSE, Pear}
with(wd, cor.test(free.sulfur.dioxide, total.sulfur.dioxide))
```

Moderate level of correlation. I was actually expecting this to be higher.

I am now going to repeat the exploration that I did for FSO2 for total FSO2.
Seeing as the two variables are correlated, and that the correlation was 
stronger between total FS02 and Quality versus free S02 and quality, this
means that it is more likely we will be able to tease out the relationship 
from our graphs.

```{r echo=FALSE, Pearso66}
total_groups <- group_by(wd, total.sulfur.dioxide)

wd.qual_by_total <- summarise(total_groups,
                             quality_mean = mean(quality),
                             quality_median = median(quality),
                             n = n())
```

This time to avoid getting caught out I'll plot the distributions of 
observations first.

```{r echo=FALSE, anulook46}
ggplot(aes(x = total.sulfur.dioxide, y = n),
       data = wd.qual_by_total) +
  geom_line() +
  coord_cartesian(xlim=c(0,160))
```

Again we see that the lower concentrations are more highly represented in the
data than the higher observations.

I have a feeling I used the wrong approach to address this earlier. Rather than 
grouping by FS02 score, I should have first created a new variable in the
original dataframe to assign each observation a percentile FSO2. THen I could
group by THAT, and the resulting graph would be adjusted for frequency.

The next couple of graphs I will remedy this mistake and also investigate
TOTAL FS02.

```{r echo=FALSE,666 }
#Use the Ntile function to group FSO2 and TSO2 by the relative rank of the score
#leaving an equal number of observations assigned to each percentile.
#This percentile will be plotted vs mean quality.

#Create new dataframes for the fixed distributions of the means.

wd2 <- wd
wd2$pfree <- ntile(wd$free.sulfur.dioxide, 
                                     n = 100)
wd2$ptotal <- ntile(wd$total.sulfur.dioxide,
                    n = 100)

free_groups_fixed <- group_by(wd2, pfree)

wd.qual_by_free_fixed <- summarise(free_groups_fixed,
                             quality_mean = mean(quality),
                             n = n())

total_groups_fixed <- group_by(wd2, ptotal)

wd.qual_by_total_fixed <- summarise(total_groups_fixed,
                             quality_mean = mean(quality),
                             n = n())
```

ok, now , lets revisit Fs02 (the pfree variable is the percentile of Fs02):


```{r echo=FALSE, 6666}
ggplot(aes(x = pfree, y = quality_mean),
       data = wd.qual_by_free_fixed ) +
  geom_point() +
  geom_smooth()
```

Well, now we can see clearly what was already confirmed in the correlation - 
there isn't really a relationship between free S02 and Quality.

Now I will repeat the processs for Total Fs02. First I will show the old way
(without adjusting for n)

```{r echo=FALSE, anulook266}
ggplot(aes(x = total.sulfur.dioxide, y = quality_mean),
       data = wd.qual_by_total) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(xlim=c(0,160))
```

As discussed before, it is difficult to extract meaning from this graph because
we know that some points are more valid than others, specifically the higher
TS02 concentraions tend to be based on less observations. There is a vague trend
downward but it isn't very steep, and it appears to start to increase!

```{r echo=FALSE, 666666}
ggplot(aes(x = ptotal, y = quality_mean),
       data = wd.qual_by_total_fixed ) +
  geom_point() +
  geom_smooth()
```

After adjusting for n and fitting our smoothedline we can now see a reasonably
clear, extremely weak negative relationship!

Not only that, but we can also see a slight drop at the lower end of the
concentration scale - could this be because more microbes were present to spoil
the wine?

I am quite pleased with the above methodology, as it can now be applied easily
to two other variables that I want to explore: Acetic acid concentration
(variable = volatile acidity) and Citric acid concentration.

Based on the descriptions of these variables, I would expect Acetic acid to be
negatively correlated with quality, and Citric acid to be positively correlated.

Acetic Acid Exploration:

```{r echo=FALSE, 77}
wd2$pacetic <- ntile(wd$volatile.acidity, 
                    n = 100)

acetic_groups<- group_by(wd2, pacetic)

wd.qual_by_acetic <- summarise(acetic_groups,
                             quality_mean = mean(quality),
                             n = n())

ggplot(aes(x = pacetic, y = quality_mean),
       data = wd.qual_by_acetic) +
  geom_point() +
  geom_smooth()
```

A nice clear, negative correlation, as we expected.

To confirm:

```{r echo=FALSE, Pearscetic}
with(wd, cor.test(volatile.acidity, quality))
```

On to Citric Acid:

```{r echo=FALSE, 777}
wd2$pcitric <- ntile(wd$citric.acid, 
                    n = 100)

citric_groups <- group_by(wd2, pcitric)

wd.qual_by_citric<- summarise(citric_groups,
                             quality_mean = mean(quality),
                             n = n())

ggplot(aes(x = pcitric, y = quality_mean),
       data = wd.qual_by_citric) +
  geom_point() +
  geom_smooth()
```

We see again a nice clear relationship that fits with the description of Citric
acid as being associated with positive qualities.
It is possible that there is a curve because the perception of 
"Freshness" by the taster is only possible above certain concentrations of 
Citric acid?

The positive correlation is confirmed below:

```{r echo=FALSE, Pearsoetic7}
with(wd, cor.test(citric.acid, quality))
```

# Bivariate Analysis

### Talk about some of the relationships you observed

The relationships I investigated were:

Free S02 versus quality

Total S02 versus quality 

Free S02 versus Total S02

Volatile acid versus quality

Citric acid versus quality.

The main result of the investigation was that Acetic acid was moderately
correlated with a decline in quality. There wasn't a relationship between free
SO2 and quality as the description of the Total S02 variable would suggest.
Confusingly however, the TS02 variable was weakly correlated with quality, and
there was a a moderate correlation between Free So2 and Total So2.

### Did you observe any interesting relationships between the other features \
###not the main feature(s) of interest)?

A strong positive correlation was observed between Free and Total SO2

### What was the strongest relationship you found?

The above.

# Multivariate Plots Section

I will start by creating a new categorical variable based on quality score.
This will allow us to combine this with our other numberic data in the same plot.

```{r echo=FALSE, Multivariate_Plot_Setup}
#Create a new categorical variable for quality to al

wd$CatQ <- cut(wd$quality, breaks = c(0, 4, 6, 8), labels = c("Poor", "Mediocre", "Good"))

#Create a new categorical variable for ACETIC 
wd$CatACETIC <- cut(wd$volatile.acidity, breaks = 3, labels = c("Low", "Medium", "High"))

#ggpairs(wd3)
```

```{r echo=FALSE, Multivariate_Plots}
wd3 <- wd[, c("volatile.acidity", "free.sulfur.dioxide", "total.sulfur.dioxide", "pH", "sulphates", "quality", "CatQ", "CatACETIC")]

ggpairs(wd3)

#ggpairs(wd3)
```
```{r echo=FALSE, Multivariate_Investigations}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide),
       data = wd3) +
       geom_point(aes(color = CatQ)) +
  geom_smooth() +
  scale_color_brewer(type = "seq")
```

As previously established, FSO2 and TS02 are linked.

We know that both distributions are skewed, however. This will impact
readability.

Lets try taking the log of both variables and see if anything jumps out to relate the
ratio of these two variables to quality:

```{r echo=FALSE, Multivariate_Investigations7}
#wd3$logTS02 <- log(wd3$total.sulfur.dioxide)
#wd3$logFS02 <- log(wd3$free.sulfur.dioxide)

ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide),
       data = wd3) +
       geom_point(aes(color = CatQ)) +
  geom_smooth() +
scale_color_brewer(type = "seq") +
  scale_x_log10(breaks = c(1, 5, 10, 15, 30)) +
  scale_y_log10()
```

The majority of the Good quality wines appear to fall below the line of best fit
- this is worth looking into further; I will come back to this in my final
graphs.

```{r echo=FALSE, Multivariate_Investigations77}
ggplot(aes(x = sulphates, y = total.sulfur.dioxide),
       data = wd3) +
       geom_point(aes(color = CatQ)) +
  geom_smooth() +
  scale_color_brewer(type = "seq")
```

I was hoping to shed some light on the interrelationship of the Sulphate
variable with TSO2, but there doesn't seem to be a clear pattern here.

Taking the log of both sulphats and TS02:

```{r echo=FALSE, Multivariate_Investigations98}
#Create a new  variable for logsulphates
#wd3$logsulphates <- log(wd3$sulphates)

ggplot(aes(x = sulphates, y = total.sulfur.dioxide),
       data = wd3) +
       geom_point(aes(color = CatQ)) +
  geom_smooth() +
  scale_color_brewer(type = "seq") +
    scale_x_log10(breaks = c(0.5, 1)) +
  scale_y_log10()
```

Lets try using the Y variable as output, in this case quality. 
Given that we have the highest correlations between quality and these two
variables.

```{r echo=FALSE, Multivariate_Investigations989}
#Create a new categorical variable for quality
ggplot(aes(x = sulphates, y = quality, colour = CatACETIC),
       data = wd3) +
       geom_jitter(alpha = 0.5, size = 0.9)
```

What about log sulfates? We know we have a skewed distribution of Sulphates,
so transforming could give us a clearer idea of what's going on:

```{r echo=FALSE, Multivariate_Investigations765}

ggplot(aes(x = sulphates, y = quality, colour = CatACETIC),
       data = wd3) +
       geom_jitter(alpha = 0.5) +
  scale_x_log10(breaks = c(0.1, 0.5, 1, 1.5))
```

This is a bit clearer. Quality has a positive association with Sulphates - we
see this reflected in that there is a slight creeping left to right as we
move up quality. The negative relationship between Volatility and quality
is also apparent in that we see more green and blue further down
the quality scale.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The ratio of mean log FSO2 to mean log TS02 appears to have a minor effect.
Noticeablymore of the high quality wines showed up at the lower end.

There did not seem to be any relationship between the amount of acetic acid
and log of sulphates.

Log TS02 and Sulphates did have a relationship - however there was a strange
plateua in the middle. I would guess that this implies that up to a certain
point, adding one increases the other - however something happens to the
sulphates at a certain concentration that prevents it being converted to S02 - 
perhaps because it is preferentially converted into something else - before the
linear relationship resumes.

### Were there any interesting or surprising interactions between features?

I was surprised that there were no high acetic acid concentrations at higher
sulfate concentrations.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x = pfree, y = quality_mean),
       data = wd.qual_by_free_fixed ) +
  geom_point() +
  geom_smooth() +
  ggtitle('Mean Quality versus FS02 Percentile') +
  ylab("Mean Quality") +
  xlab("Free Sulphur Dioxide (Percentile Rank)")

```

### Description One

This graph establishes that despite the information provided there isn't a
relationship between FS02 and Quality.

### Plot Two
```{r echo=FALSE, Plot_Two}
library(RColorBrewer)

ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide),
       data = wd3) +
       geom_jitter(aes(color = CatQ)) +
    scale_color_brewer(type = 'seq',
                     guide = guide_legend(title = 'Colour', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) + 
  geom_smooth() +
    ggtitle('TSO2 Versus FS02') +
  xlab("Log Free Sulphur Dioxide") +
  ylab("Log Total Sulphur Dioxide") +
  scale_x_log10(breaks = c(1, 10, 50)) +
  scale_y_log10()
```

### Description Two

By omitting the mediocre wines we can see more clearly whether the good ones
fall on the high or low side of the F/T So2 ratio. In this case it is still
ambiguous however it does look like more good wines fall under the line.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = sulphates, y = total.sulfur.dioxide),
       data = wd3) +
       geom_point(aes(color = CatQ)) +
  geom_smooth() +
    ggtitle('Sulphates to S02 by Quality') +
  xlab("Log Free Sulphates") +
  ylab("Log Total Sulphur Dioxide") +
      scale_color_brewer(type = 'seq',
                     guide = guide_legend(title = 'Colour', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +
    scale_x_log10(breaks = c(0.5, 1, 2)) +
  scale_y_log10()


```

### Description Three
By carrying out the same modification we can take out some of the noise and
examine the relationship between the ratio of TS02 to Sulphates by Quality.

The graph shows little clear relationship between the variables, though there
is a cluster of poor quality wines at low levels of both Sulphates and Sulphur
Dioxide.
------

# Reflection

Overall I would say that despite the lack of a continuous output variable
and my overall unfamiliarity with wine I was able to explore several
interesting aspects of the data, in particular testing what appear to be
inaccurate assumptions about how several variables would relate to quality
based on the descriptions provided.

The major hurdle in this project was grappling with the inherent biases
of the distributions and finding a clear way of accounting for this.

My background reading on the topic revealed that wine is an incredibly complex
product, so there are many many more chemical variables to explore.
Moreover as mentioned previously, the output metric - quality is very
questionable given the evidence against the ability of wine tasters to 
consistently grade wines. However it should be noted that the quality metric
used in this dataset is itself a median, and the gradings are rather broad, so
this would offset any inaccuracy and still provide a good general sense of 
whether a wine is good or not.
